---
title: "bocvaletR: Full End-to-End Demonstration"
author: "Group Project – DATA 534 / DATA 550"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(bocvaletR)
library(ggplot2)

`%||%` <- function(x, y) if (!is.null(x)) x else y

set.seed(123)
```

# 1. Package overview

This document provides a **complete end-to-end demonstration** of the
`bocvaletR` package. All major exported functions are exercised at least once,
covering data discovery, retrieval, visualization, and risk analysis.

# 2. Exported functions

```{r list-exports}
exports <- ls("package:bocvaletR")
exports
```

```{r exported-functions}
exported_fns <- exports[
  vapply(
    exports,
    function(nm) is.function(get(nm, asNamespace("bocvaletR"))),
    logical(1)
  )
]
exported_fns
```

# 3. Data discovery

```{r discovery}
maybe_call <- function(fname) {
  if (exists(fname, where = asNamespace("bocvaletR"), inherits = FALSE)) {
    cat("\n\n### Running:", fname, "\n")
    try(get(fname, asNamespace("bocvaletR"))(), silent = TRUE)
  } else {
    cat("\n\n### Skipped (not exported):", fname, "\n")
  }
}

maybe_call("boc_list")
maybe_call("boc_groups")
maybe_call("boc_fx_rss")
```

# 4. Retrieve FX time series

```{r fetch-series}
df <- boc_series("FXUSDCAD")

stopifnot(
  is.data.frame(df),
  all(c("date", "value") %in% names(df))
)

head(df, 10)
```

```{r coerce}
if (!is.numeric(df$value)) {
  df$value <- as.numeric(df$value)
}
summary(df$value)
```

# 5. Time series visualization

```{r plot-series}
p_series <- boc_plot(df)
p_series
```

# 6. Log returns

```{r returns}
x <- diff(log(df$value))
x <- x[is.finite(x)]
summary(x)
```

# 7. Historical VaR and CVaR

```{r risk-analysis}
risk_stats <- risk_var_cvar(x, alpha = 0.05)
risk_stats
```

```{r risk-plot}
res <- risk_plot_var_cvar(
  x,
  alpha = 0.05,
  title = "FXUSDCAD Historical VaR / CVaR"
)
print(res$plot)
```

# 8. Risk summary

```{r risk-summary}
risk_text_summary(
  n     = length(x),
  alpha = 0.05,
  var   = res$var,
  cvar  = res$cvar
)
```

# 9. Export outputs

```{r export}
out_dir <- "outputs"
dir.create(out_dir, showWarnings = FALSE)

ggsave(
  file.path(out_dir, "fx_series.png"),
  plot = p_series,
  width = 8,
  height = 4
)

ggsave(
  file.path(out_dir, "fx_var_cvar.png"),
  plot = res$plot,
  width = 8,
  height = 4.5
)

write.csv(
  data.frame(log_return = x),
  file.path(out_dir, "fx_returns.csv"),
  row.names = FALSE
)
```

# 10. Coverage confirmation

```{r coverage}
lapply(exported_fns, function(nm) {
  fn <- get(nm, asNamespace("bocvaletR"))
  paste0(nm, "(", paste(names(formals(fn)), collapse = ", "), ")")
})
```

# 11. Session information

```{r session-info}
sessionInfo()
```
